### 2013-12-03 manuscripts/reply/short-reply.tex
Working on empirical Bayes section of `short-reply.tex`.

\citet{Hickerson2013} suggest that \citet{Oaks2012} should have used their
data to inform their prior on divergence times (i.e., an empirical Bayesian
approach).
However, \citet{Oaks2012} did use empirically informed priors to mimic
empirical situations in which a large amount of prior information is available;
they did so to assess the prior sensitivity of the method and to determine if
uniform priors in such situations are narrow enough to avoid the bias toward
models with few divergence events.
The results of \citet{Oaks2012} and \citet{Hickerson2013} showed the method is
highly sensitive to the prior on divergence times.
Furthermore, \citet{Oaks2012} found the method remained biased toward clustered
divergences even under the informed priors.
The main argument of \citet{Hickerson2013} is that \citet{Oaks2012} did not
go far enough, and suggest a very narrow, highly informed uniform prior
on divergence times is necessary to avoid the method's preference for
models with few divergence-time parameters.
Such an empirical Bayesian approach to model selection raises some theoretical
and practical concerns, some of which were discussed by \citet{Oaks2012} (see
the last paragraph of ``Assessing prior sensitivity of \msb'' in
\citet{Oaks2012}); we expand on this here.

@@

For our re-analysis of the Philippines dataset we used the model-averaging
approach of \citet{Hickerson2013}, but with a reduced set of prior models to
avoid mixing time units.
We used five prior models, all of which had priors on population sizes of
$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
U(0.0001, 0.05)$.
Following \citet{Hickerson2013}, each of these models had the following
priors on divergence time parameters:
$M_1$, $\divt{} \sim U(0, 0.1)$;
$M_2$, $\divt{} \sim U(0, 1)$;
$M_3$, $\divt{} \sim U(0, 5)$;
$M_4$, $\divt{} \sim U(0, 10)$; and
$M_5$, $\divt{} \sim U(0, 20)$.
We simulated $1\e6$ random samples from each of the models for a total of
$5\e6$ prior samples.
For each model, we retained the 10,000 samples with the smallest Euclidean
distance from the observed summary statistics, standardizing the statistics
using the prior means and standard deviations of the given model.
From the remaining 50,000 samples, we then retained the 10,000 samples with the
smallest Euclidean distance from the observed summary statistics, this time
standardizing the statistics using the prior means and standard deviations
across all five models.
We then repeated this analysis twice, replacing the $M_1$ with
$M_{1A}$ and $M_{1B}$, which differ only by having priors on divergence
times of $\divt{} \sim U(0, 0.01)$ and $\divt{} \sim U(0, 0.001)$,
respectively.
While we suspect the prior of $\divt{} \sim U(0, 0.1)$ used by
\citet{Hickerson2013} likely excludes the true divergence times of at least
some of the 22 taxa, we are nearly certain that these narrower priors are
incorrect and exclude most, if not all, of the divergence times of the
Philippine taxa.

Our results show that the model-averaging approach of \citet{Hickerson2013}
does not reduce the method's bias toward models with less parameter space,
but rather allows it to manifest in a rather dangerous way.
The method strongly prefers the prior model with the narrowest distribution on
divergence times across all three of our analyses, even when this model is
almost certainly incorrect and excludes the true divergence times of the
Philippine taxa (Table~\ref{tabModelChoiceEmpirical}).
Unfortunately, ``checking'' the priors by plotting the summary statistics from
1000 random samples of each prior model along the first two orthogonal axes of
a principle component analysis, as recommended by \citet{Hickerson2013},
provides no warning of a problem (Figure~\ref{figPCA}).
Given that the results of \citet{Hickerson2013} strongly prefer the models with
the narrowest prior on divergence times, it seems quite likely that their
model-averaged results are dominated by models that exclude at least some of
the 22 true divergence times, making their results difficult to interpret
(in addition to the error of time units).

>>>>

\citet{Hickerson2013} suggest that \citet{Oaks2012} should have used their
data to inform their prior on divergence times (i.e., an empirical Bayesian
approach).
However, \citet{Oaks2012} did use empirically informed priors to mimic
empirical situations in which a large amount of prior information is available;
they did so to assess the prior sensitivity of the method and to determine if
uniform priors in such situations are narrow enough to avoid the bias toward
models with few divergence events.
The results of \citet{Oaks2012} and \citet{Hickerson2013} showed the method is
highly sensitive to the prior on divergence times.
Furthermore, \citet{Oaks2012} found the method remained biased toward clustered
divergences even under the informed priors.

The main argument of \citet{Hickerson2013} is that the priors used by
\citet{Oaks2012} were not informative enough.
They suggest a very narrow, highly informed uniform prior on divergence times
is necessary to avoid the method's preference for models with few
divergence-time parameters.
Such an empirical Bayesian approach to model selection raises some theoretical
and practical concerns, some of which were discussed by \citet{Oaks2012} (see
the last paragraph of ``Assessing prior sensitivity of \msb'' in
\citet{Oaks2012}); we expand on this here.

@@

There was an error in the methods of \citet{Hickerson2013} resulting in
an erroneous model that mixes different units of time (see SI for
details).
For our re-analysis of the Philippines dataset we followed the model-averaging
approach of \citet{Hickerson2013}, but with a reduced set of prior models to
avoid the error.
We used five prior models, all of which had priors on population sizes of
$\meanDescendantTheta{} \sim U(0.0001, 0.1)$ and $\ancestralTheta{} \sim
U(0.0001, 0.05)$.
Following \citet{Hickerson2013}, each of these models had the following
priors on divergence time parameters:
$M_1$, $\divt{} \sim U(0, 0.1)$;
$M_2$, $\divt{} \sim U(0, 1)$;
$M_3$, $\divt{} \sim U(0, 5)$;
$M_4$, $\divt{} \sim U(0, 10)$; and
$M_5$, $\divt{} \sim U(0, 20)$.
We simulated $1\e6$ random samples from each of the models for a total of
$5\e6$ prior samples.
For each model, we retained the 10,000 samples with the smallest Euclidean
distance from the observed summary statistics, standardizing the statistics
using the prior means and standard deviations of the given model.
From the remaining 50,000 samples, we then retained the 10,000 samples with the
smallest Euclidean distance from the observed summary statistics, this time
standardizing the statistics using the prior means and standard deviations
across all five models.
We then repeated this analysis twice, replacing the $M_1$ with
$M_{1A}$ and $M_{1B}$, which differ only by having priors on divergence
times of $\divt{} \sim U(0, 0.01)$ and $\divt{} \sim U(0, 0.001)$,
respectively.
While we suspect the prior of $\divt{} \sim U(0, 0.1)$ used by
\citet{Hickerson2013} likely excludes the true divergence times of at least
some of the 22 taxa, we are nearly certain that these narrower priors are
incorrect and exclude most, if not all, of the divergence times of the
Philippine taxa.

Our results show that the model-averaging approach of \citet{Hickerson2013}
does not reduce the method's bias toward models with less parameter space,
but rather allows it to manifest in a rather dangerous way.
The method strongly prefers the prior model with the narrowest distribution on
divergence times across all three of our analyses, even when this model is
almost certainly incorrect and excludes the true divergence times of the
Philippine taxa (Table~\ref{tabModelChoiceEmpirical}).

\citet{Hickerson2013} vetted the priors used in their model-averaging approach
via ``graphical checks,'' in which the summary statistics from 1000 random
samples of each prior model are plotted along the first two orthogonal axes of
a principle component analysis (see Figure 1 of \citet{Hickerson2013}).
To determine if such prior predictive analyses would indicate the $M_{1A}$ and
$M_{1B}$ are are problematic, we performed these graphical checks on our prior
models.
Unfortunately, these prior predictive checks provide no warning that these
priors are too narrow (Figure~\ref{figPCA}).
Rather, they suggest these incorrect priors are ``better fit''
(Figure~\ref{figPCA}A--C) than the valid priors used by \citet{Oaks2012}
(Figure~\ref{figPCA}D--F).

Given that the results of \citet{Hickerson2013} strongly prefer the models with
the narrowest prior on divergence times, it seems quite likely that their
model-averaged results are dominated by models that exclude at least some of
the 22 true divergence times, making their results difficult to interpret
(in addition to the error of time units).
